Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2017-01-25 manuscript/manuscript_pve_editor.Rmd
started response to editor.  big change is dropping asat figure and adding it to table 3

```{r, echo=FALSE}
allomfig <- "1"
airvarsfig <- "2"
allocfig <- "3"
asatfig <- "4"
achemfig <- "5"
cuefig <- "6"

@@

\
Leaf dark respiration rates (*R*) were measured on each seedling on the same dates as AC~i~ curves. Freshly detached leaves were collected at least 1 hour after sundown and placed inside a conifer chamber attached to the Licor 6400. Measurements were taken at a reference [CO~2~] of 400 $\mu$l l^-1^ while leaf temperature was maintained at current ambient conditions. *R* was logged 5 times for each leaf after CO~2~ and water vapour flux values stabilized in the conifer chamber. Reported values of *R* are standardized rates at 25 &deg;C using a *Q*~10~ value (1.86) developed for these seedlings in a separate experiment (Drake *et al.* unpublished). Leaf area and dry mass were recorded for each leaf during gas exchange campaigns.

@@

##Leaf, root and soil chemistry
Each leaf used in gas exchange measurements and subsamples of harvested roots were oven dried at 60&degC to a constant mass and ball milled for analysis of nitrogen (N) content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were passed through a 1 mm sieve to remove organic material, air dried at 40&degC for 72 h and ball milled for analysis of N. Nitrogen concentrations of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer and the percentage of N in samples was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the standard Vienna Pee Dee Belemnite scale.

@@

\
Total seedling biomass at harvest was significantly different across container volumes (*P* < 0.001) and between container treatments and Free seedlings (*P* < 0.001, Table `r pvetab1`). On average, harvested biomass of Free seedlings was 84 % higher than that of seedlings in containers. Plant biomass was positively correlated with total leaf area across all treatments (*R*^2^ = 0.97, *P* < 0.001). Differences in biomass partitioning to leaves, stems and roots were not different across soil volumes after variation in seedling biomass across treatments was accounted for in the analysis (Figure `r allocfig`a,b). The final harvest root:shoot biomass ratio was conserved across all treatments, with a slightly higher shoot than root mass ($\overline{x}$ = 0.904, 95 % CI = [0.846,1.119]). The ratio of leaf to fine root mass was also not different between treatments (Figure `r allocfig`c).

@@

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the Free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ (Figure `r asatfig`) and *A*~max~ (Table `r pvetab3`) were consistently higher in Free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ with *A*~max~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was highly correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~max~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~max~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~max~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

\
Both *J*~max~ and *V*~cmax~ were significantly higher in Free seedlings (30 % and 26 %, respectively) than container-grown seedlings with little variation between container volume treatments (Table `r pvetab3`). Overall, the *J*~max~ / *V*~cmax~ ratio was conserved across all treatments (1.6&pm;0.02), which is consistent with findings across many tree species [@medlyn1996optimal; @leuning1997scaling; @medlyn1999effects; @warren2003photosynthesis; @crous2008elevated]. Leaf dark respiration rates were not significantly different across soil volumes (Table `r pvetab3`). The *g*~1~ parameter, generated for each seedling from the Medlyn *et al.* [-@medlyn2011reconciling] optimal stomatal conductance model, was lowest in the Free seedling treatment and was marginally different across soil volume treatments (Table `r pvetab3`).

@@

#Discussion
This study utilized a novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling growth. Our goal was to identify how sink limitation of growth modified the C balance of seedlings. We found a strong reduction of growth in containerized plants. Leaf *A*~n~ was also reduced, but this reduction was insufficient to fully explain the reduction in growth, implying that components such as whole-plant respiratory loss were also likely affected.

@@

\
**Table 3**. Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment. *A*~max~, *R*~dark~ and *g*~s~ are each measured at 25 &deg;C. Values of *A*~max~, *g*~s~ and *g*~1~ represent overall means across measurement campaigns (n=6). R, *J*~max~ and *V*~cmax~ values are means of two measurement campaigns at beginning and end of gas exchange measurements. Different letters represent significant differences between treatments. The volume effect *P* value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

```{r table3, echo=FALSE, warning=FALSE}
dust(table3)%>%
  sprinkle_colnames("Volume (L)", 
                        "*A*~max~ ($\\mu$mol m^-2^ s^-1^)", 
                        "*R*~dark~ ($\\mu$mol m^-2^ s^-1^)",
                        "*J*~max~ ($\\mu$mol m^-2^ s^-1^)",
                        "*V*~cmax~ ($\\mu$mol m^-2^ s^-1^)",
                        "*g*~s~ (mol m^-2^ s^-1^)",
                        "*g*~1~") %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

@@

```{r allometry, fig.width=6, fig.height=7.8, echo=FALSE,  dev="png", dpi=600,eval=includeFigs}
source("master_scripts/allometryplotting.R")
**Figure 1**. Seedling growth over the duration of the experiment was strongly reduced in containerized seedlings compared to freely rooted seedlings. Height growth (a), diameter growth (b), and seedling leaf area estimated from leaf counts (c) measured weekly on *Eucalyptus tereticornis* seedlings across the experiment duration in 2013. Shown are soil volume treatment means &pm; standard error.
\
```{r airvars, fig.width=6, fig.height=8, echo=FALSE, dev="png", dpi=600,eval=includeFigs}
source("master_scripts/airvars_plotting.R")
**Figure 2**. Environmental drivers over the course of the experiment. Daily maximum and minimum temperature (a), total daily PPFD (b), and daily maximum vapour pressure deficit (c) in 2013. 
\
```{r allocation, fig.width=6, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE, dev="png", dpi=600,eval=includeFigs}
source("master_scripts/allocation_plotting.R")
**Figure 3**. Biomass partitioning at final harvest did not respond strongly to soil volume treatment. (a) Relative partitioning to leaves, stems, and roots at the final harvest, shown as treatment means (standard errors omitted for clarity). (b) Bi-variate relationships between mass partitioning to leaves and stems + roots. Each symbol is a plant. Lines represent standardized major axis fitting of the log-transformed allometric relationships of leaf mass fraction by treatment. (c) Total plant leaf mass as a function of fine root biomass with &pm; standard error. The dashed line is a 1:1 line and the solid line represents the significant log-log model fit (*R*^2^ = 0.82) with equation: log(x) = -0.36 + 0.983 * log(y).
\
```{r asat, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE, message=FALSE, dev="png", dpi=600,eval=includeFigs}
source("master_scripts/A_stats_plotting.R")
**Figure 4**. Light-saturated photosynthesis (A~sat~) was reduced in containerized seedlings compared to freely-rooted seedlings ('Free'). Shown are soil volume treatment means &pm; standard error, across all measurement campaigns (n = 6), of light saturated rates of photosynthesis at 25 &deg;C. Different letters represent significant differences between treatments (p < 0.05).
\
```{r A_tncnitro, fig.width=8, fig.height=6, echo=FALSE,warning=FALSE, message=FALSE, dev="png", dpi=600, eval=includeFigs}
source("master_scripts/Amax_TNC_N_plotting2.R")
**Figure 5**. Photosynthetic capacity on a leaf mass basis (A~mass~, measured at saturating light and [CO~2~]) was correlated to both leaf starch and nitrogen content. (a) A~mass~ was negatively related to leaf starch (shown as percentage of dry mass), in particular for leaves with higher nitrogen content (N~mass~). (b) A~mass~ was positively related to N~mass~ (expressed as N per unit mass after subtracting TNC content). In both panels, each symbol is an individual measurement of A~mass~ (data across all campaigns are shown). Colors represent bins levels (n = 5) of both leaf starch and nitrogen grouped from low to high. Lines represents predictions, for each bin level, from the linear mixed effects model equation of A~mass~ as a function of starch and nitrogen. The marginal *R*^2^ (fixed effects only) was 0.37 and the conditional *R*^2^ (fixed and random effects) was 0.48 for the complete model. The equation for the full model is y = - 221.96 + 28.58(N) - 3.31(Starch) - 0.87(N*Starch).
\
```{r cuefig, fig.width=6, fig.height=6, echo=FALSE, dev="png", dpi=600, eval=includeFigs}
source("master_scripts/cuefig.R")
**Figure 6**. The final biomass attained per unit total photosynthesis over the duration of the experiment. Note the higher ratio for the freely rooted seedlings compared to the containerized seedlings, indicating a higher carbon-use efficiency. Shown are treatment means, as the model scaling instantaneous leaf photosynthesis to total plant photosynthesis was run once for every treatment.

>>>>

```{r, echo=FALSE}
allomfig <- "1"
airvarsfig <- "2"
allocfig <- "3"
achemfig <- "4"
cuefig <- "5"

@@

\
Leaf dark respiration rates (*R*) were measured on each seedling on the same dates as AC~i~ curves. Freshly detached leaves were collected at least 1 hour after sundown and placed inside a conifer chamber attached to the Licor 6400. Measurements were taken at a reference [CO~2~] of 400 $\mu$l l^-1^ while leaf temperature was maintained at current ambient conditions (11.5-13.4 &deg;C). *R* was logged 5 times for each leaf after CO~2~ and water vapour flux values stabilized in the conifer chamber. Reported values of *R* are standardized rates at 25 &deg;C using a *Q*~10~ value (1.86) developed for these seedlings in a separate experiment (Drake *et al.* unpublished). Leaf area and dry mass were recorded for each leaf during gas exchange campaigns.

@@

##Leaf, root and soil chemistry
Each leaf used in gas exchange measurements and subsamples of harvested roots were oven dried at 60 &deg;C to a constant mass and ball milled for analysis of nitrogen (N) content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were passed through a 1 mm sieve to remove organic material, air dried at 40 &deg;C for 72 h and ball milled for analysis of N. Nitrogen concentrations of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer and the percentage of N in samples was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the standard Vienna Pee Dee Belemnite scale.

@@

\
Total seedling biomass at harvest was significantly different across container volumes (*P* < 0.001) and between container treatments and Free seedlings (*P* < 0.001, Table `r pvetab1`). On average, harvested biomass of Free seedlings was 84 % higher than that of seedlings in containers. Plant biomass was positively correlated with total leaf area across all treatments (*R*^2^ = 0.97, *P* < 0.001). Differences in biomass partitioning to leaves, stems and roots were not different across soil volumes after variation in seedling biomass across treatments was accounted for in the analysis (Figure `r allocfig`a,b). The final harvest root:shoot biomass ratio was conserved across all treatments, with a slightly higher shoot than root mass ($\overline{x}$ = 0.90, 95 % CI = [0.85,1.12]). The ratio of leaf to fine root mass was also not different between treatments (Figure `r allocfig`c).

@@

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the Free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ and *A*~max~ (Table `r pvetab3`) were consistently higher in Free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~max~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~max~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~max~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

\
Both *J*~max~ and *V*~cmax~ were significantly higher in Free seedlings (30 % and 26 %, respectively) than container-grown seedlings with little variation between container volume treatments (Table `r pvetab3`). Overall, the *J*~max~ / *V*~cmax~ ratio was conserved across all treatments (1.6&pm;0.02). Leaf dark respiration rates were not significantly different across soil volumes (Table `r pvetab3`). The *g*~1~ parameter, generated for each seedling from the Medlyn *et al.* [-@medlyn2011reconciling] optimal stomatal conductance model, was lowest in the Free seedling treatment and was marginally different across soil volume treatments (Table `r pvetab3`).

@@

#Discussion
This study utilized a novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling growth. Our goal was to identify how sink limitation of growth modified the C balance of seedlings. We found a strong reduction of growth in containerized plants. Leaf *A*~n~ was also reduced, but this reduction was insufficient to fully explain the reduction in growth, implying that components such as non-leaf respiratory losses were also likely affected.

@@

\
**Table 3**. Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment. *A*~max~, *A*~sat~, *R*~dark~ and *g*~s~ are each measured at 25 &deg;C. Values of *A*~max~, *A*~sat~, *g*~s~ and *g*~1~ represent overall means across measurement campaigns (n=6). R, *J*~max~ and *V*~cmax~ values are means of two measurement campaigns at beginning and end of gas exchange measurements. Different letters represent significant differences between treatments. The volume effect *P* value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

```{r table3, echo=FALSE, warning=FALSE}
dust(table3)%>%
  sprinkle_colnames("Volume (L)", 
                        "*A*~max~ ($\\mu$mol m^-2^ s^-1^)", 
                        "*A*~sat~ ($\\mu$mol m^-2^ s^-1^)",
                        "*R*~dark~ ($\\mu$mol m^-2^ s^-1^)",
                        "*J*~max~ ($\\mu$mol m^-2^ s^-1^)",
                        "*V*~cmax~ ($\\mu$mol m^-2^ s^-1^)",
                        "*g*~s~ (mol m^-2^ s^-1^)",
                        "*g*~1~") %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

@@

```{r allometry, fig.width=6, fig.height=7.8, echo=FALSE,  dev="png", dpi=600,eval=includeFigs}
source("master_scripts/allometryplotting.R")
```  
**Figure 1**. Seedling growth over the duration of the experiment was strongly reduced in containerized seedlings compared to freely rooted seedlings. Height growth (a), diameter growth (b), and seedling leaf area estimated from leaf counts (c) measured weekly on *Eucalyptus tereticornis* seedlings across the experiment duration in 2013. Shown are soil volume treatment means &pm; standard error.
\
```{r airvars, fig.width=6, fig.height=8, echo=FALSE, dev="png", dpi=600,eval=includeFigs}
source("master_scripts/airvars_plotting.R")
```  
**Figure 2**. Environmental drivers over the course of the experiment. Daily maximum and minimum temperature (a), total daily PPFD (b), and daily maximum vapour pressure deficit (c) in 2013. 
\
```{r allocation, fig.width=6, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE, dev="png", dpi=600,eval=includeFigs}
source("master_scripts/allocation_plotting.R")
```  
**Figure 3**. Biomass partitioning at final harvest did not respond strongly to soil volume treatment. (a) Relative partitioning to leaves, stems, and roots at the final harvest, shown as treatment means (standard errors omitted for clarity). (b) Bi-variate relationships between mass partitioning to leaves and stems + roots. Each symbol is a plant. Lines represent standardized major axis fitting of the log-transformed allometric relationships of leaf mass fraction by treatment. (c) Total plant leaf mass as a function of fine root biomass with &pm; standard error. The dashed line is a 1:1 line and the solid line represents the significant log-log model fit (*R*^2^ = 0.82) with equation: log(x) = -0.36 + 0.983 * log(y).
\
```{r A_tncnitro, fig.width=8, fig.height=6, echo=FALSE,warning=FALSE, message=FALSE, dev="png", dpi=600, eval=includeFigs}
source("master_scripts/Amax_TNC_N_plotting2.R")
```  
**Figure 4**. Photosynthetic capacity on a leaf mass basis (A~mass~, measured at saturating light and [CO~2~]) was correlated to both leaf starch and nitrogen content. (a) A~mass~ was negatively related to leaf starch (shown as percentage of dry mass), in particular for leaves with higher nitrogen content (N~mass~). (b) A~mass~ was positively related to N~mass~ (expressed as N per unit mass after subtracting TNC content). In both panels, each symbol is an individual measurement of A~mass~ (data across all campaigns are shown). Colors represent bins levels (n = 5) of both leaf starch and nitrogen grouped from low to high. Lines represents predictions, for each bin level, from the linear mixed effects model equation of A~mass~ as a function of starch and nitrogen. The marginal *R*^2^ (fixed effects only) was 0.37 and the conditional *R*^2^ (fixed and random effects) was 0.48 for the complete model. The equation for the full model is y = - 221.96 + 28.58(N) - 3.31(Starch) - 0.87(N*Starch).
\
```{r cuefig, fig.width=6, fig.height=6, echo=FALSE, dev="png", dpi=600, eval=includeFigs}
source("master_scripts/cuefig.R")
```  
**Figure 5**. The final biomass attained per unit total photosynthesis over the duration of the experiment. Note the higher ratio for the freely rooted seedlings compared to the containerized seedlings, indicating a higher carbon-use efficiency. Shown are treatment means, as the model scaling instantaneous leaf photosynthesis to total plant photosynthesis was run once for every treatment.
